---
title: "Problem Set 6 R Solutions"
author: Matthew Alampay Davis
date: April 13, 2021
output:
  pdf_document: default
---

```{r, include = FALSE}
setwd('~/Documents/Grad School/Columbia/Y2/Intro Metrics TA/Past problem sets/PS7')
library(readstata13)
library(dplyr)
library(magrittr)
library(estimatr)
library(lmtest)
library(sandwich)
library(ggplot2)
library(knitr)
# library(AER)
library(gmm)
library(ivmodel)
opts_chunk$set(tidy.opts = list(width.cutoff=60), tidy=TRUE)
select <- dplyr::select
```

# Load the data

```{r}
daron <- read.dta13('maketable4.dta')
```

# Question 1

## 1a

Filter out countries without a European colonial history or that are missing data

```{r}
daron %<>% filter(baseco==1)
```

Columns 1-2

To do 2SLS, let's use *iv_robust*, which like *lm_robust* comes from the *estimatr* package. Note: for some reason, Stata uses a different type of robust standard error when estimating IV models so the argument for **se_type here is 'HC0' rather than 'HC1' (which is the same as 'stata')**.

The regression formula is as usual except with a bar "|". Regressors go on the left side of the bar and instruments on the right.

```{r}
reg1 <- iv_robust(logpgp95 ~ avexpr | logem4, data = daron, se_type = 'HC0') # different SEs
summary(reg1)
```

Now the second model. Note here that we want to model *lat_abst* as exogenous (i.e. we only want to instrument for *avexpr*). So we include *lat_abst* on both sides of the bar "|" in the formula argument. If we hadn't included it on the RHS, then the model would assume *lat_abst* was an endogenous variable and we'd have more endogenous variables than instruments, which would prevent the regression from running.

```{r}
reg2 <- iv_robust(logpgp95 ~ lat_abst + avexpr | lat_abst + logem4, data = daron, se_type = 'HC0')
summary(reg2)
```

For regressions 1 and 2, no standard errors differ by more than 0.20

## 1b

### 1b i) Method 1

This question has you do the first-stage and second-stage regressions yourself.

```{r}
# First stage
mod.1bi.1 <- lm_robust(avexpr ~ logem4, data = daron, se_type = 'stata')
# Second stage
mod.1bi.2 <- lm_robust(daron$logpgp95 ~ mod.1bi.1$fitted.values, se_type = 'stata')
summary(mod.1bi.2)
```

Note that the solutions tell us the standard errors that come out of this exercise are incorrect. So even though they're wrong, I used 'stata' standard errors above so they match what's in the Stata solutions. It seems like whenever Stata runs non-IV regressions, they use 'stata' standard errors and when they run IV in one command, they use 'HC0' standard errors. As far as I know, there's no deep reason for this.

Following the solutions, you have to go back to using the all-in-one-step method from Part A using 'HC0' standard errors to get the proper standard errors:

```{r}
mod.1bii <- iv_robust(logpgp95 ~ avexpr | logem4, data = daron, se_type = 'HC0')
summary(mod.1bii)
```

### 1b ii) Method 2

```{r}
cov(select(daron, logpgp95, avexpr, logem4))
```

### 1b iii) Method 3

```{r}
# First stage regression
mod.1biii.1 <- lm_robust(avexpr ~ logem4, data = daron, se_type = 'stata')
# Second stage regression
mod.1biii.2 <- lm_robust(logpgp95 ~ logem4, data = daron, se_type = 'stata')
# Divide the coefficient in the second by the coefficient in the first
coef(mod.1biii.2)[2]/coef(mod.1biii.1)[2]

```


$$
\beta_3 = \frac{-0.9067352}{-0.9602404}\approx 0.9443
$$

## 1c

Repeat for regressions 3 and 4, excluding the four countries with exceptional GDPs

```{r}
daron2 <- filter(daron, rich4 != 1)
reg3 <- iv_robust(logpgp95 ~ avexpr | logem4, data = daron2, se_type = 'HC0')
summary(reg3)
reg4 <- iv_robust(logpgp95 ~ avexpr + lat_abst | logem4 + lat_abst, data = daron2, se_type = 'HC0')
summary(reg4)
```

Repeat for regressions 5 and 6, excluding the African countries

```{r}
daron3 <- filter(daron, africa != 1)
reg5 <- iv_robust(logpgp95 ~ avexpr | logem4, data = daron3, se_type = 'HC0')
reg6 <- iv_robust(logpgp95 ~ avexpr + lat_abst | logem4 + lat_abst, data = daron3, se_type = 'HC0')
```

For regressions 3 through 6, two standard errors differ by more than 0.20: t > he SE on latitude in the first stage in regression 4 and in regression 6.

## 1d

No exclusions. Add continent indicators as controls

Create a variable called "other" that equals one if the observation is not in the base group, not coded Africa or Asia, and not in the Americas:

```{r}
daron %<>% mutate(other = shortnam %in% c('AUS', 'NZL', 'MLT'))
table(daron$other)
```

Then the regression:

```{r}
reg7 <- iv_robust(logpgp95 ~ avexpr + africa + asia + other | logem4 + africa + asia + other,
                  data = daron,
                  se_type = 'HC0')
summary(reg7)

reg8 <- iv_robust(logpgp95 ~ avexpr + africa + asia + other + lat_abst | logem4 + africa + asia + other + lat_abst,
                  data = daron,
                  se_type = 'HC0')
summary(reg8)

reg9 <- iv_robust(loghjypl ~ avexpr | logem4,
                  data = daron,
                  se_type = 'HC0')
summary(reg9)
```

## 1e

Any recap, highlights, points of interest, opinions, or assertions gets points.

## 1f

We'll modify Regression 9

IV-GMM requires the *gmm* package

```{r}
y <- daron$loghjypl # outcome
d <- daron$avexpr # endogenous regressor
z <- daron$logem4 # instrument

iv.gmm <- gmm(y ~ d, x = z)
summary(iv.gmm)
```

IV-LIML requires the *ivmodel* package

```{r}


iv.liml <- ivmodel(y, d, z)
LIML(iv.liml) %>% unlist
```

# Question 2

Does not require R so see the official solutions