---
title: "PS8 R Solutions"
author: "Matthew Alampay Davis"
date: "December 12, 2021"
output:
  pdf_document: default
---

```{r, include = FALSE}
setwd('~/Dropbox/Grad School/Columbia/Y3/Metrics TA/Make-up final')
library(readstata13)
library(Quandl)
library(lubridate)
library(dplyr)
library(magrittr)
library(sandwich)
library(ggplot2)

library(estimatr)
library(car)
library(lmtest)
library(knitr)
```

# Question 1

## Part a

```{r}
fish <- read.dta13('fishdata.dta')
fish.mod1 <- lm_robust(qty ~ price, se_type = 'stata', data = fish)
fish.mod2 <- lm_robust(qty ~ price + day1 + day2 + day3 + day4 + cold + rainy, se_type = 'stata', data = fish)
fish.mod3 <- iv_robust(qty ~ price | stormy, se_type = 'HC0', data = fish, diagnostics = T)
fish.mod4 <- iv_robust(qty ~ price + day1 + day2 + day3 + day4 + cold + rainy | stormy + day1 + day2 + day3 + day4 + cold + rainy, se_type = 'HC0', data = fish, diagnostics = T)
```

## Part b

```{r}
fish.mod1$coefficients['price']
fish.mod3$coefficients['price']
```

## Part c

```{r}
fish.mod3.manual1 <- lm_robust(price ~ stormy, se_type = 'stata', data = fish)
fish.mod3.manual2 <- lm_robust(qty ~ fish.mod3.manual1$fitted.values, se_type = 'stata', data = fish)
summary(fish.mod3.manual2)
fish.mod4.manual1 <- lm_robust(price ~ stormy + day1 + day2 + day3 + day4 + cold + rainy, se_type = 'stata', data = fish)
fish.mod4.manual2 <- lm_robust(qty ~ fish.mod4.manual1$fitted.values + day1 + day2 + day3 + day4 + cold + rainy, se_type = 'stata', data = fish)
```

## Part d

Yes for coefficients, no for standard errors.

## Part e

If it's relevant and exogenous

## Part f

Relevance in regression 3

```{r}
fish.diagnostics3 <- iv_robust(qty ~ price | stormy, se_type = 'HC0', data = fish, diagnostics = T)
fish.diagnostics3$diagnostic_first_stage_fstatistic
```

F statistic is 21.042 so we can reject the null hypothesis stormy is not relevant.

## Part g

```{r}
fish.ivs1 <- iv_robust(qty ~ price | stormy + windspd + windspd2, se_type = 'HC0', data = fish, diagnostics = T)
fish.ivs2 <- iv_robust(qty ~ price + day1 + day2 + day3 + day4 + cold + rainy | stormy + windspd + windspd2 + day1 + day2 + day3 + day4 + cold + rainy, se_type = 'HC0', data = fish, diagnostics = T)
```

## Part h

If demand is unit elastic, then the coefficient on price is equal to -1

```{r}
stat.1 <- (-1.095+1)/0.4060
stat.2 <- (-1.11337+1)/0.4473
```

The t-statistics or on the order of -0.23 to -0.25 so we cannot reject the null hypohtesis of unit elasticity in either case

## Part i

```{r}
summary(fish.ivs1)
summary(fish.ivs2)
```

In both cases, the F statistics are under 10 so the instruments do appear to be weak. This suggests our IV estimate of the demand elasticity is not valid.

## Part j

For column 4 (i.e. back to using just one instrument), we have

```{r}
summary(fish.mod4)
```

Since we have one instrument for one potentially endogenous regressor, we cannot perform a test of overidentifying restrictions (it is not overidentified). In the enhanced model where we include windspd and windspd2 as instruments, we do have 3 instruments for one endogenous regressor and we get the following:

```{r}
summary(fish.ivs2)
```

The J-statistic is 1.287 which has a corresponding p-value of 0.525. Thus we cannot reject the null hypothesis that all instruments are exogenous.

# Question 2

Non-empirical

# Question 3

```{r}
macro <- read.dta13('macro_us.dta') %>%
  mutate(rgdp.l1 = lag(rgdp),
         rgdp.l2 = lag(rgdp, n = 2),
         rgdp.l3 = lag(rgdp, n = 3),
         rgdp.l4 = lag(rgdp, n = 4),
         ip.l1 = lag(ip, n = 1))
```

## Part a

```{r}
m.trunc <- ceiling(0.75*nrow(macro)^(1/3))
adl <- lm(rgdp ~ rgdp.l1 + rgdp.l2 + rgdp.l3 + rgdp.l4 + ip.l1, data = macro)
nw.se <- NeweyWest(adl, lag = m.trunc, prewhite = F, adjust = T)
coeftest(adl, vcoov = nw.se)
```
## Part b

```{r}
tail(macro)
fore <- 36.764695 + 0.828997*19478.89 + 0.121238*19368.31 + 0.021034*19055.65 + 0.026991*18767.78 + 0.953835*99.8782
fore
```

# Question 4

Non-empirical

# Question 5

```{r}
airfare <- read.dta13('airfare.dta')
```

## Part a

## Part b

## Part c

## Part d

## Part e

## Part f

## Part g

## Part h

## Part i

## Part j